---
title: "Writing Research Reports"
author: "Dr. Christian Lucero ©"
date: "Summer II 2019"
output:
  pdf_document:
    highlight: haddock
    keep_tex: yes
    latex_engine: xelatex
    number_sections: yes
    # toc: yes
header-includes:
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{wrapfig}
- \usepackage{float}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{xcolor}
geometry: margin = 0.5in
editor_options:
  chunk_output_type: console
urlcolor: blue
---




```{r setup, include=FALSE}
# This is the setup chunk
#  Here you can set global options for the entire document
library(knitr) # I recommend doing this

# Although you can call functions from a library using the following notation
#  without loading the entire library.
knitr::opts_chunk$set(echo = TRUE, 
                      comment = NA, # Recommended
                      fig.path = "./figures/",  # Store all figures here in relative path
                      fig.align = "center",
                      fig.width = 7,
                      fig.height = 7,
                      message = FALSE, # Turn off load messages
                      warning = FALSE # Turn off warnings,
                      )
```

\clearpage

```{r include=FALSE}

# You should set your working directory at the very beginning of your R Markdown file
setwd("C:/Users/marco/Desktop/CMDA 4654/Projects/Project 1")

# You should load all of the libraries that you plan to use here unless you plan
#  on using the notation library_name::function_name()

library(tidyverse) 
library(pander)  # We used pander here, but some alternative libraries are below:
library(xtable)
library(kableExtra)
```


# Abstract

It is important that you learn how to write a proper research report.  In this document we will show a few examples of the type of output that we expect to see in a well-written and professional looking report.


# Introduction and General Guidelines

One of the most important things to keep in mind is that a professional written report is not a homework assignment.  You are expected to present a document that does not include unnecessary computer code and output that cannot be handled by a more general audience.  This does not mean that you should not be showing plots and output from the computer, they just need to be well-formatted.

In this document we will briefly show a few examples of how to present some of your work while adhering to a few guidelines:

1. R code should not be seen in the main body unless it's a report specifically involving the use of R.  Even then only include the **important** snippets.  Save all other code for the appendix.

2. All plots and tables should have clear and easy text and other plot elements (this should be true for your homework as well!).  Make sure that the axis text, and annotations can be read if you were to print things out on a regular sheet of paper.  It's okay to make a plot or table bigger and rotate it, if necessary.

3. All plots and tables should have a title or caption.  In general they should have a Figure or Table number so that they can be easily referenced in the text.

4. Never make a plot or table unless you plan on discussing it in the text.

5. Do not show the raw output from R. The results of statistical analyses should be in a tables.

6. Use sections and subsections to properly organize your work.

7. Reports rendered directly to a .pdf using LaTeX.  Keep in mind, these are supposed to be reports that may be printed out.  Rendering to .html and then converting to .pdf can result in a document that looks awkward when printed and may have many formatting issues.

\vspace{.2in}
\hrule
\vspace{.2in}

Your document should include the following sections:

1. Introduction

2. Description of the Data 
    * You should briefly describe your data.  Where did you get it?  What does it correspond to?  What are some of the important features?

3. Methods
    * You should generally have a brief section that discusses the methods you used.  You don't have to go into great detail, but it's important that we understand why you chose to use the methods that you used.  You can provide some brief analysis here as part of your justifications.

4. Results and Discussion
    * You should include the results of your analyses appropriately labeled and into subsections as needed.  All results, including tables, plots, etc., need a discussion.

5. Conclusions 
    * What are the ultimate take away messages that you want to leave the reader with based upon your analysis?  If this was a recommendation to a boss or client, what would that recommendation be?

6. Citations
    * Biblography section.